Causal Forest Approach for Site-specific Input Management via On-farm Precision Experimentation

### 2022-04-19 Writing/manuscript_cea.rmd
large updates for 3rd round submission

---
title: "Causal Forest Approach for Site-specific Input Management via On-farm Precision Experimentation"
author: Shunkei Kakimoto^[Corresponding author, Department of Agricultural Economics, University of Nebraska-Lincoln 102 Filley Hall 1625 Arbor Drive Lincoln, NE 68583, USA, E-mail address; skakimoto3@hunskers.unl.edu], Taro Mieno^[Department of Agricultural Economics, University of Nebraska-Lincoln, Lincoln, NE 68583, USA], Takashi S. T. Tanaka^[Artificial Intelligence Advanced Research Center, Faculty of Applied Biological Sciences, Gifu University, Gifu 501-1193, Japan], and David S. Bullock^[Department of Agricultural and Consumer Economics, University of Illinois, Urbana, IL 61801, USA]
output:
  bookdown::word_document2:
    fig_caption: yes
    number_sections: yes
    global_numbering: true
    # reference_docx: word_template.docx
    reference_docx: word_template_format.docx
    # pandoc_args: ["-Fpandoc-crossref"]
bibliography: ML_VRA.bib
csl: computers-and-electronics-in-agriculture.csl
eqnPrefix: 
      - "Eq"
abstract: "Estimating site-specific crop yield response to changes in input (e.g., seed, fertilizer) management is a critical step in making economically optimal site-specific input recommendations. Past studies have attempted to estimate yield response functions using various Machine Learning (ML) methods, including the Random Forest (RF), Boosted Random Forest (BRF), and Convolutional Neural Network (CNN) methods. This study proposes use of the Causal Forest (CF) model, which is one of the emerging ML methods that comprise \"Causal Machine Learning.\" Unlike previous yield-prediction-oriented ML methods, CF focuses strictly on estimating heterogeneous treatment effects (changes in yields that result from changes in input application rates) of inputs. We report results of using Monte Carlo simulations assuming various production scenarios to test the effectiveness of CF in estimating site-specific economically optimal nitrogen rates (EONRs), comparing CF with the yield-prediction-oriented ML methods RF, BRF, and CNN. CF's estimations of site-specific EONRs were superior under all scenarios considered. Furthermore, we also show that the quality of a model's yield predictions provides little if any information about the quality of its EONR predictions."  
---

@@

Information about crop yield response input management is critical for the identification of site-specific economically optimal input rates (EOIRs) [@bullock1998does], and the difficulty in estimating site-specific yield response functions accurately within a field is the primary explanation of the persistent relatively low the adoption rate of variable rate crop input application technology [@lowenberg2019setting]. Historically, site-specific yield response functions have been estimated using the data generated from strip trials conducted on only parts of fields [@makowski2001statistical; @scharf2002corn; @bullock2002adding; @mamo2003spatial; @hurley2004estimating; @anselin2004spatial; @scharf2005field; @ruffo2006site; @Miao2006; @puntel2019development].
But recently whole-field on-farm precision experimentation (OFPE) [@bullock2019] has allowed implementation of improved statistical experimental designs (e.g., Latin Square, Randomized Block) on entire fields [@laurent_kyveryga_makowski_miguez_2019; @licht_witt_2019]. Data generated from whole-field OFPEs typically have greater local variation in experimental input application rates and soil/field characteristics than do strip trial designs, which has improved the statistical identification of heterogeneous impacts of input management changes on yields. The resultant availability of more spatially detailed experiment data has encouraged a resurgence of interest in estimating site-specific yield response functions. For example, @barbosa2020modeling, @barbosa2020risk, @krause2020random and @Gardner21 used OFPE data to estimate site-specific yield response functions and economically optimal site-specific N fertilizer and/or seed rates. These recent studies used machine learning (ML) methods extensively in place of more traditional parametric modeling approaches. @barbosa2020modeling and @barbosa2020risk applied the Convolutional Neural Network approach, @krause2020random used Random Forest-based approaches, and @Gardner21 used the shape-constrained generalized additive model. 
But this trend of applying ML methods may raises concerns. Historically, agronomists have worked meticulously to develop clever experimental designs to identify causes of yield response to input management. But the aforementioned recent applications of ML methodologies pay little attention to identify such causes, instead attempting to validate their models by estimating yield prediction accuracy. But the ultimate goal of OFPE is the accurate prediction of EOIRs, not yields. These are distinct objectives, and achievement of one need not imply the other.

The lack of and need for causal interpretation of the results of ML research applications have been increasingly recognized, in particular in economics and medicine [@athey2017beyond; @athey2018impact; @Athey2019a; @moraffah2020causal; @Storm2020; @arti2020research; @scholkopf2021toward], and causal machine learning (CML) methods were recently developed in response. Unlike traditional prediction-oriented ML methods, CML focuses on identifying causal impacts of an event (in our context, of an increase or decrease in the input application rate). The growing academic literature applying CML methods includes @richens2020improving, who's medical research pointed out that the inability of the traditional ML methods to identify cause-and-effect relationships between disease and patients' characteristics could lead to misdiagnoses, and showed that CML-based diagnoses could improve the accuracy of diagnostic prediction. In addition, @carbo2020machine used a CML to identify characteristics causing bank customers to adopt online banking services, and @bozorgi2020process applied CML to study the mining processes.

In this paper, we use the Causal Forest (CF) approach developed by @athey2016recursive and @Wager2018a to examine the accuracy of methods that estimate site-specific economically optimal N fertilizer rates (EONRs). We compare the EONR prediction performances of our CF-based approach to the widely used ML methods Random Forest (RF), Boosted Random Forest (BRF), and Convolutional Neural Network (CNN), and we examine how yield prediction performance is related to EONR prediction performance, which allow us to examine whether model selection should be based on yield prediction accuracy when the ultimate goal is EOIR identification. We use Monte Carlo simulations in which we synthetically generate yield and soil/field characteristics datasets to create stylized "field." Because real-world "true" EOIRs are unobservable, real-world data cannot be used to test the accuracy of statistical methods estimating EOIRs. This important difference contrasts starkly with yield prediction, since actual yields can be ground-truthed.

@@

In the following, we describe the fundamental information necessary to identify EOIR in general: the site-specific treatment effect of input application on yield. We then discuss the causal forest method in detail, show how it is particularly suitable to obtain this fundamental information, and contrast it to more traditional prediction-oriented ML methods, such as RF, BRF, and CNN.

@@

Recent studies using ML methods to estimate the effectiveness of various agricultural input management strategies first estimate yield function, $f(I, \mathbf{c})$ (e.g., @barbosa2020modeling; @barbosa2020risk; @krause2020random; @Gardner21) and then derive changes in yield associated with changes in input rate to estimate site-specific input recommendations. On the other hand, Causal Forest is designed specifically to estimate a discrete version of Eq. \@ref(eq:eqn2) directly, examining the discrete changes in yield that result from discrete *changes* in the input application rate $I$, bypassing the estimation of $f(I, \mathbf{c})$. It may be suspected that CF will outperform other ML methods that identify EOIR indirectly through yield prediction. We now details how CF works below and why CF may perform better in predicting EOIR than other prediction-oriented ML methods used in the previous studies.

@@

CF was developed specifically for identifying heterogeneous treatment effects [@athey2016recursive; @Wager2018a]. Like RF, CF works by building an ensemble of trees, called *causal trees* (CTs), by recursively partitioning covariate space into two leaves, where each leaf estimates conditional average treatment effects as the mean differences between the dependent variable (here, yield) of the treated and control observations that belong to the leaf (see Appendix A for an example of a CT that can help further clarify how CF works).

While both RF and CF build trees by recursively partitioning samples, they differ critically in their criteria for splitting. In RF, the sample is split to minimize the mean squared error (MSE) of yield prediction in building trees. Unfortunately, the MSE of treatment effect predictions is not useful for building a CT since the true treatment effects are never observable. However, @athey2016recursive showed that minimizing the expected MSE of the treatment effect is equivalent to maximizing the variance of treatment effects across the resulting two leaves and minimizing the within-leaf variance. Consequently, by splitting samples in a way that maximizes the variance of treatment effects across the resulting two leaves, CF overcomes the fundamental problem created by the unobservability of true treatment effects. CF requires that the tree-building process is "honest" for treatment effects estimation to be unbiased. Honest tree-building first randomly splits in two the training data samples, then uses one of the subsamples to determine how to split the tree, and uses the other to estimate the treatment effects [@athey2016recursive; @Wager2018a]. This honest subsampling technique avoids overestimation of the heterogeneity of estimated treatment effects. Since CF splits samples to maximize the heterogeneity of estimated treatment effects, using the same sample for both the sample splitting and treatment effect calculations renders itself sensitive to noise or outliers, allowing high noise levels to exaggerate the treatment effects.

@@

Whether the indirect (yield-prediction-first) and direct approach is used can lead to meaningful statistical differences when it comes to EOIR estimation. The conventional prediction-oriented ML methods are designed to predict yield, $y_i(I,\mathbf{c})$, well, but not necessarily treatment effects. Since the loss function used in prediction-oriented ML uses (typically squared) residuals in yield prediction, prediction-oriented ML is naturally optimized for predicting the "level" of yield instead of identifying the causal impact of the input on yield. CF, on the other hand, is designed specifically to estimate the causal impact of input. Indeed, @Wager2018a used Monte Carlo simulations to show that CF provided more accurate estimates of heterogeneous treatment effects than those provided by RF. However, the underlying data generating process of their simulations was extremely simple, with the treatment effect influenced by only a single variable. CF has not been tested for more complicated data generating processes, such as those represented by yield response functions in which the treatment effect of an input can be affected by multiple variables in a non-linear fashion. Moreover, CF has never been compared to more advanced ML methods such as BRF and CNN that are considered to have prediction capabilities greater than RF's. The study reported here builds on @Wager2018a to compare CF to these prediction-oriented ML methods (not just RF, but also BRF and CNN) in more realistic agronomic experiment settings.

@@

Monte Carlo simulations were conducted for nitrogen experiments for corn production to estimate site-specific EONRs. The simulations' first step generated one thousand synthetic fields that "resembled" an actual field. Field characteristics and yield were simulated assuming the Mitscherlich-Baule (MB) yield response function. We used the RF, BRF, and CNN ML methods to estimate yield response functions, and we used the CF method to estimate treatment effects. For each method, we estimated site-specific EONRs using the training dataset, and checked the statistical performance of the predicted site-specific EONR using the validation dataset. Detailed descriptions of these steps are provided below. The R and python codes that implement the simulation analyses are publicly accessible as a Mendeley dataset [@Kakimoto2022codes], or as a Github repository at https://github.com/Shunkei3/VRA_with_CF.git.

@@

The simulated field was created based on an actual 51-ha Illinois field on which the Data-Intensive Farm Management (DIFM) project [@bullock2019] has run multiple OFPEs. Figure \@ref(fig:field-map) illustrates that simulated trials consisted of 384 18.3 m $\times$ 73.2 m "plots," each of which was assigned an N fertilizer application rate. Each plot was made up of four (4-rows $\times$ 1-column) 18.3 m $\times$ 18.3 m "subplots," which were unit of analysis used in subsequent statistical analysis. Each subplot contained of a 6-rows $\times$ 6-columns grid of thirty-six 3.05 m $\times$ 3.05 m "cells." All field characteristics and yield data were generated at the cell level.

@@

The first term on the right-hand side of Eq. \@ref(eq:eqn3) is the deterministic yield component. The variable $N$ represents the nitrogen fertilizer application rate (kg/ha) which the trial designs varied by plot. Also in Eq. \@ref(eq:eqn3), $\alpha$, $\beta$, and $ymax$ are spatially varying parameters that govern how yield level is determined in conjunction with the applied N rate. They reflect local cell-level production conditions, such as soil characteristics and topography. For example, $\alpha$, which represents what the yield would be when no nitrogen is applied in conjunction with $ymax$, may reflect soil organic matter and residual nitrogen from previous year's production. $ymax$, which determines the maximum yield attainable, may include other micro nutrients like phosphorus and soil depth. $\beta$ can be interpreted as reflecting how efficiently the crop uses N to create grain mass, and might depend on soil properties that affect the availability of applied N fertilizer to the crop, say due to N immobilization, leaching, denitrification, and volatilization [@johnson2005nitrogen; @alva2006nitrogen]. Figure \@ref(fig:vis-MB-curve) presents example visualization of site-specific yield response functions for some sites.

@@

The error term of the yield $\varepsilon$ represents the composite of all the unobserved factors that affect crop yield beyond $\alpha$, $\beta$, $ymax$ and $N$. Each cell $(i,j)$'s "true" optimal N application rate, denoted $N^{opt}_{i,j}$, was calculated by solving a profit maximization problem by cell assuming that the prices of corn and N fertilizer were $\$0.138/kg$ and $\$1.323/kg$, respectively. These prices are used consistently for our simulations. See Appendix C.1 for further details on how the parameters were generated and sample maps of the parameters and true economically optimal N rates.

@@

In each simulation round, five field-trial N rates were determined based on the range of the round's cell-specific EONRs. 
Let $Q_q$ ($q \in [0,1]$) denote the $100 \times q$ percent quantile of the round's cell-specific EONRs. Then, the five trial design rates were $N_1 = Q_{0.05} - 20$, $N_5 = Q_{0.95} + 20$, $N_2 = N_1 + 0.25(N_5 - N_1)$, $N_3 = N_1 + 0.50(N_5 - N_1)$, $N_4 = N_1 + 0.75(N_5 - N_1)$. 
In each round, a Latin square design was followed to assign each plot one of the five N rates so established. The left-most panel of Figure \@ref(fig:fig-nExp-yield) shows an example of a trial design. The N rate was orthogonal to the error term in every simulation's example of trial design. This is important because if they were correlated, bias in the estimation of the impact of N on yield would be introduced into all the ML methods. Since the treatment effect at each leaf is obtained simply by taking the difference of the average value of the dependent variable between the control and treated group within the leaf, CF requires the exogeneity of the treatment conditional on the independent variables to estimate the treatment effect consistently. Note that the violation of this condition would lead to biased estimation using the CF and the other ML methods.

A cell-level dataset for the whole field, $\{(y_{i,j}, N^{opt}_{i,j}, N_{i,j}, \mathbf{c}_{i,j}): i \in \{1,2,\dots,1536\}, j \in \{1,2,\dots,36\}\}$ was generated by plugging in the parameter values, the error term, and the assigned nitrogen rates into Eq. \@ref(eq:eqn3). The cell-level data were then aggregated up to the subplot level by taking the mean of the cell-level values inside each subplot, which were the unit of observation. The subplot data were therefore

@@

To evaluate the performances of RF, BRF, CNN and CF, we examined four modeling scenarios, each defined by a dataset assumed to be available to the researcher. The first modeling scenario, denoted "aby" assumed that the researcher has a dataset of the actual values $\alpha_{i}$, $\beta_{i}$, $ymax_{i}$ in the yield response function for every subplot $i$ and use $\alpha$, $\beta$ and $ymax$ as covariates in the ML models. This scenario represents an ideal situation for the researcher, in which all the relevant variables are observed, but the true functional form of heterogeneous yield response functions are not known. More realistic scenarios can be compared with this ideal scenario.

In subsequent modeling scenarios, we made the simulated experimental data more realistic by varying the degree to which the researcher understood what the needed covariates' values were those in the dataset. In the second scenario, denoted "abytt," we introduced additional variables $\theta^1$ and $\theta^2$, and the researcher's dataset contained values of $\alpha_{i}$, $\beta_{i}$, $ymax_{i}$, $\theta^1_{i}$ and $\theta^2_{i}$ for all subplots $i=1,\dots,1536$. $\theta^1$ and $\theta^2$ had no effect on yield, but they were spatially correlated with $\beta_{i}$. This scenario was meant to reflect the lack of scientific consensus about which variables explain heterogeneous yield response to N, making it likely that the researcher would include irrelevant variables like $\theta^1$ and $\theta^2$ in real-world data analysis.

In the third scenario, denoted "aabbyy," the researcher's dataset did not contain values of $\alpha_{i}$, $\beta_{i}$, or $ymax_{i}$. Rather, it contained subplot-level values of permutations of $\alpha$, $\beta$, and $ymax$. Each of the three parameters were split into two parts in a spatially correlated manner, and six new covariates were created (see Appendix C.2 for the details about this parameter-splitting process). These were $\alpha^{1}_{i}$, $\alpha^{2}_{i}$, $\beta^{1}_{i}$, $\beta^{2}_{i}$, $ymax^{1}_{i}$, and $ymax^{2}_{i}$. The two split variables were spatially correlated to reflect that many of the soil/field characteristics used to analyze OFPE data are spatially correlated. The values of $\alpha_{i}$, $\beta_{i}$ and $ymax_{i}$ did not indicate specific soil or field characteristics, but were functions of such. In reality, scientists cannot directly observe variables that can accurately predict yield plateau level, soil N content and N uptake efficiency. Rather, they use multiple observed soil/field characteristics to explain such phenological phenomena. For example, instead of using $ymax_{i}$ directly, some topographical land features and soil properties such as elevation and soil sand content could be used as yield-limiting factors [@jiang2004effect; @kravchenko2000correlation]. In reality, scientists may include more than three variables as covariates. This scenario should reduce the accuracy of EONR modeling for all the models compared to the ideal case.

@@

The RF, BRF, and CNN methods all followed the same conceptual steps to estimate site-specific EONR: 1) estimate the yield production functions and 2) calculate site-specific EONRs based on that estimated function. For the given modeling scenario, RF, BRF and CNN predict yield by using all the available covariates as explanatory variables in the estimation process. The `grf` package (version 1.2.0 [@tibshirani2018package]) in R was used for RF and BRF modeling. RF ann BRF from the `grf` package have several tunable hyperparameters (e.g., the minimum node size in each tree, the number of covariates used for node splitting, and parameters involving an honest tree-building process), and the optimal values for these parameters were selected by cross-validation. Individual trees in a forest were built with randomly subsampled observations, leading to a different prediction in every forest even when the same dataset was used. To reduce the variance of predictions for prediction accuracy, we set the number of trees (*num.trees*) at a high value, $2000$.

>>>>

---
title: "Causal Forest Approach for Site-specific Input Management via On-farm Precision Experimentation"
author: Shunkei Kakimoto^[Corresponding author, Department of Agricultural Economics, University of Nebraska-Lincoln 102 Filley Hall 1625 Arbor Drive Lincoln, NE 68583, USA, E-mail address; skakimoto3@hunskers.unl.edu], Taro Mieno^[Department of Agricultural Economics, University of Nebraska-Lincoln, Lincoln, NE 68583, USA], Takashi S. T. Tanaka^[Artificial Intelligence Advanced Research Center, Faculty of Applied Biological Sciences, Gifu University, Gifu 501-1193, Japan], and David S. Bullock^[Department of Agricultural and Consumer Economics, University of Illinois, Urbana, IL 61801, USA]
output:
  bookdown::word_document2:
    fig_caption: yes
    number_sections: yes
    global_numbering: true
    # reference_docx: word_template.docx
    reference_docx: word_template_format.docx
    # pandoc_args: ["-Fpandoc-crossref"]
bibliography: ML_VRA.bib
csl: computers-and-electronics-in-agriculture.csl
eqnPrefix: 
      - "Eq"
abstract: "Estimating site-specific crop yield response to changes to input (e.g., seed, fertilizer) management is a critical step in making economically optimal site-specific input management recommendations. Past studies have attempted to estimate yield response functions using various Machine Learning (ML) methods, including the Random Forest (RF), Boosted Random Forest (BRF), and Convolutional Neural Network (CNN) methods. This study proposes use of the Causal Forest (CF) model, which is one of the emerging ML methods that comprise \"Causal Machine Learning.\" Unlike previous yield-prediction-oriented ML methods, CF focuses strictly on estimating heterogeneous treatment effects (changes in yields that result from changes in input application rates) of inputs. We report results of using Monte Carlo simulations assuming various production scenarios to test the effectiveness of CF in estimating site-specific economically optimal nitrogen rates (EONRs), comparing CF with the yield-prediction-oriented ML methods RF, BRF, and CNN. CF's estimations of site-specific EONRs were superior under all scenarios considered. We also show that the quality of a model's yield predictions provides little if any information about the quality of its EONR predictions."  
---

@@

Information about crop yield response input management is critical for the identification of site-specific economically optimal input rates (EOIRs) [@bullock1998does], and the difficulty in estimating site-specific yield response functions accurately within a field is the primary explanation of the persistent relatively low the adoption rate of variable rate crop input application technology [@lowenberg2019setting]. Historically, site-specific yield response functions have been estimated using the data generated from strip trials conducted on only parts of fields [@anselin2004spatial; @bullock2002adding; @hurley2004estimating; @makowski2001statistical; @mamo2003spatial; @Miao2006; @puntel2019development; @ruffo2006site; @scharf2002corn; @scharf2005field].
But recently whole-field on-farm precision experimentation (OFPE) [@bullock2019] has allowed implementation of improved statistical experimental designs (e.g., Latin Square, Randomized Block) on entire fields [@laurent_kyveryga_makowski_miguez_2019; @licht_witt_2019]. Data generated from whole-field OFPEs typically have greater local variation in experimental input application rates and soil/field characteristics than do strip trial designs, which has improved the statistical identification of heterogeneous impacts of input management changes on yields. The resultant availability of more spatially detailed experiment data has encouraged a resurgence of interest in estimating site-specific yield response functions. For example, Barbosa et al. [-@barbosa2020risk; -@barbosa2020modeling], @krause2020random and @Gardner21 used OFPE data to estimate site-specific yield response functions and economically optimal site-specific N fertilizer and/or seed rates. These recent studies used machine learning (ML) methods extensively in place of more traditional parametric modeling approaches. Barbosa et al. [-@barbosa2020risk; -@barbosa2020modeling] took a Convolutional Neural Network approach, @krause2020random used Random Forest-based approaches, and @Gardner21 used the shape-constrained generalized additive model. 
But this trend of applying ML methods may raise concerns. Historically, agronomists have worked meticulously to develop clever experimental designs to identify causes of yield response to input management. But the aforementioned recent applications of ML methodologies pay little attention to identifying such causes, but instead attempt to validate their models by estimating yield prediction accuracy. But the ultimate goal of OFPE is the accurate prediction of EOIRs, not yields. These are distinct objectives, and achievement of one need not imply the other.

The lack of and need for causal interpretation of the results of ML research applications have been increasingly recognized, in particular in economics and medicine [@athey2017beyond; @athey2018impact; @Athey2019a; @moraffah2020causal; @Storm2020; @arti2020research; @scholkopf2021toward], and causal machine learning (CML) methods were recently developed in response. Unlike traditional prediction-oriented ML methods, CML focuses on identifying causal impacts of an event (in our context, of an increase or decrease in the input application rate). The growing academic literature applying CML methods includes @richens2020improving, whose medical research pointed out that the inability of the traditional ML methods to identify cause-and-effect relationships between disease and patients' characteristics could lead to misdiagnoses, and showed that CML-based diagnoses could improve the accuracy of diagnostic prediction. In addition, @carbo2020machine used a CML to identify characteristics causing bank customers to adopt online banking services, and @bozorgi2020process applied CML to study mining processes.

We use the Causal Forest (CF) approach developed by @athey2016recursive and @Wager2018a to examine the accuracy of methods that estimate site-specific economically optimal N fertilizer rates (EONRs). We compare the EONR prediction performances of our CF-based approach to the widely used ML methods Random Forest (RF), Boosted Random Forest (BRF), and Convolutional Neural Network (CNN), and we examine the relation between yield prediction performance and EONR prediction performance, which allows us to study whether model selection should be based on yield prediction accuracy when the ultimate goal is EOIR identification. We use Monte Carlo simulations in which we synthetically generate yield and soil/field characteristics datasets to create stylized "field." Because real-world "true" EOIRs are unobservable, real-world data cannot be used to test the accuracy of statistical methods estimating EOIRs. This important difference contrasts starkly with yield prediction, since actual yields can be ground-truthed.

@@

In the following, we describe the fundamental information necessary to identify EOIR in general: the site-specific treatment effect of input application on yield. We then discuss the causal forest method in detail, show how it is particularly suited to the gathering of this fundamental information, and contrast it to more traditional prediction-oriented ML methods, such as RF, BRF, and CNN.

@@

Recent studies using ML methods to estimate the effectiveness of various agricultural input management strategies first estimate yield function, $f(I, \mathbf{c})$ (e.g., Barbosa et al. [-@barbosa2020risk; -@barbosa2020modeling]; @krause2020random; and @Gardner21) and then derive changes in yield associated with changes in the input rate to estimate site-specific input application recommendations. On the other hand, Causal Forest is designed specifically to estimate a discrete version of Eq. \@ref(eq:eqn2) directly, examining the discrete changes in yield that result from discrete *changes* in the input application rate $I$, bypassing the estimation of $f(I, \mathbf{c})$. It may be suspected that CF will outperform other ML methods that identify EOIR indirectly through yield prediction. Below we offer details about how CF works and may perform better in predicting EOIR than other prediction-oriented ML methods previously reported.

@@

CF was developed specifically for identifying heterogeneous treatment effects [@athey2016recursive; @Wager2018a]. Like RF, CF works by building an ensemble of trees, called *causal trees* (CTs), by recursively partitioning covariate space into two leaves, where each leaf estimates conditional average treatment effects as the mean differences between the dependent variable (here yield) of the treated and control observations that belong to the leaf. (See Appendix A for an example.)

While both RF and CF build trees by recursively partitioning samples, they differ critically in their criteria for splitting. In RF, the sample is split to minimize the mean squared error (MSE) of yield prediction in building trees. Unfortunately, the MSE of treatment effect predictions is not useful for building a CT since the true treatment effects are never observable. However, @athey2016recursive showed that minimizing the expected MSE of the treatment effect is equivalent to maximizing the variance of treatment effects across the resulting two leaves and minimizing the within-leaf variance. Consequently, by splitting samples in a way that maximizes the variance of treatment effects across the resulting two leaves, CF overcomes the fundamental problem created by the unobservability of true treatment effects. CF requires that the tree-building process be "honest" for treatment effects estimation to be unbiased. Honest tree-building first randomly splits the training data samples in two, then uses one of the subsamples to determine how to split the tree, and the other to estimate the treatment effects [@athey2016recursive; @Wager2018a]. This honest subsampling technique avoids overestimation of the heterogeneity of estimated treatment effects. Since CF splits samples to maximize the heterogeneity of estimated treatment effects, using the same sample for both the sample splitting and treatment effect calculations renders itself sensitive to noise or outliers, allowing high noise levels to exaggerate the treatment effects.

@@

Whether the indirect (yield-prediction-first) or direct approach is used can lead to meaningful statistical differences when it comes to EOIR estimation. The conventional prediction-oriented ML methods are designed to predict yield $y_i(I,\mathbf{c})$ well, but not necessarily treatment effects. Since the loss function used in prediction-oriented ML uses (typically squared) residuals in yield prediction, prediction-oriented ML is naturally optimized for predicting the "level" of yield instead of identifying this causal impact of the input on yield. CF, on the other hand, is designed specifically to estimate this causal impact. Indeed, Wager and Athey’s[-@Wager2018a] Monte Carlo simulations showed that the estimates of heterogeneous treatment effects provided by CF were more accurate than those provided by RF. However, their simulations' underlying data-generating process was extremely simple, with the treatment effect influenced by only a single variable. CF has not been tested for more complicated data generating processes, such as those represented by yield response functions in which the treatment effect of an input can be affected by multiple variables in a non-linear fashion. Moreover, CF has never been compared to more advanced ML methods such as BRF and CNN that are considered to have prediction capabilities greater than RF's. The study reported here builds on @Wager2018a to compare CF to these prediction-oriented ML methods (not just RF, but also BRF and CNN) in more realistic agronomic experiment settings.

@@

Estimates of site-specific EONRs were conducted using Monte Carlo simulations of nitrogen fertilizer application experiments for corn production. The simulations' first step generated one thousand synthetic fields that "resembled" an actual field. Field characteristics and yield were simulated assuming the Mitscherlich-Baule (MB) yield response function. We used the RF, BRF, and CNN ML methods to estimate yield response functions, and we used the CF method to estimate treatment effects. For each method, we estimated site-specific EONRs using the training dataset, and checked the statistical performance of the predicted site-specific EONR using the validation dataset. Detailed descriptions of these steps are provided below. The R and Python codes that implement the simulation analyses are publicly accessible as a Mendeley dataset [@Kakimoto2022codes], or as a Github repository at https://github.com/Shunkei3/VRA_with_CF.git.

@@

The simulated field was created based on an actual 51-ha Illinois field on which the Data-Intensive Farm Management (DIFM) project [@bullock2019] has run multiple OFPEs. Figure \@ref(fig:field-map) illustrates that simulated trials consisted of 384 18.3 m $\times$ 73.2 m "plots," each of which was assigned an N fertilizer application rate. Each plot was made up of four (4-rows $\times$ 1-column) 18.3 m $\times$ 18.3 m "subplots," which were unit of analysis used in subsequent statistical analysis. Each subplot contained a 6-row $\times$ 6-column grid of thirty-six 3.05 m $\times$ 3.05 m "cells." All field characteristics and yield data were generated at the cell level.

@@

The first term on the right-hand side of Eq. \@ref(eq:eqn3) is the deterministic yield component. The variable $N$ represents the nitrogen fertilizer application rate (kg/ha) which the trial designs varied by plot. Also in Eq. \@ref(eq:eqn3), $\alpha$, $\beta$, and $ymax$ are spatially varying parameters that govern how yield level is determined in conjunction with the applied N rate. They reflect local cell-level production conditions, such as soil characteristics and topography. For example, $\alpha$, which represents what the yield would be when no nitrogen is applied in conjunction with $ymax$, may reflect soil organic matter and residual nitrogen from previous year's production. $ymax$, which determines the maximum yield attainable, may include other micro nutrients like phosphorus and soil depth. $\beta$ can be interpreted as reflecting how efficiently the crop uses N to create grain mass, and might depend on soil properties that affect the availability of applied N fertilizer to the crop, say due to N immobilization, leaching, denitrification, and volatilization [@johnson2005nitrogen; @alva2006nitrogen]. Figure \@ref(fig:vis-MB-curve) presents examples of site-specific yield response functions for five sites.

@@

The yield error term $\varepsilon$ represents the impact of the composite of all the unobserved factors that affect crop yield that are not accounted for by $\alpha$, $\beta$, $ymax$ and $N$. Each cell $(i,j)$'s "true" optimal N application rate, denoted $N^{opt}_{i,j}$, was calculated by solving a profit maximization problem by cell assuming that the prices of corn and N fertilizer were $\$0.138/kg$ and $\$1.323/kg$ in all simulations. See Appendix C.1 for further details on how the parameters were generated and sample maps of the parameters and true economically optimal N rates.

@@

In each simulation round, five field-trial N rates were determined based on the range of the round's cell-specific EONRs. 
Let $Q_q$ ($q \in [0,1]$) denote the $100 \times q$ percentile of the round's cell-specific EONRs. Then, the five trial design rates were $N_1 = Q_{0.05} - 20$, $N_5 = Q_{0.95} + 20$, $N_2 = N_1 + 0.25(N_5 - N_1)$, $N_3 = N_1 + 0.50(N_5 - N_1)$, $N_4 = N_1 + 0.75(N_5 - N_1)$. 
In each round, a Latin square design was followed to assign to each plot one of the five N rates so established. The left-most panel of Figure \@ref(fig:fig-nExp-yield) shows an example of a trial design. The N rate was orthogonal to the error term in every simulation's trial design. This is important because correlation between N and the error term would introduce bias in the estimation of the impact of N on yield for all ML methods. Since the treatment effect at each leaf is obtained simply by taking the difference of the average value of the dependent variable between the control and treated group within the leaf, to estimate the treatment effect consistently CF requires the exogeneity of the treatment conditional on the independent variables. Note that violation of this condition would lead to biased estimation using the CF and the other ML methods.

A cell-level dataset for the whole field, $\{(y_{i,j}, N^{opt}_{i,j}, N_{i,j}, \mathbf{c}_{i,j}): i \in \{1,2,\dots,1536\}, j \in \{1,2,\dots,36\}\}$ was generated by plugging in the parameter values, the error term, and the assigned nitrogen rates into Eq. \@ref(eq:eqn3). The cell-level data were then aggregated up to the subplot level by taking the mean of the cell-level values inside each subplot, which were the unit of observation of the generated dataset:

@@

To evaluate the performances of RF, BRF, CNN and CF, we examined four modeling scenarios, each defined by a dataset assumed to be available to the researcher. The first modeling scenario, denoted "aby" assumed that the researcher has a dataset of the actual values of yield function parameters $\alpha_{i}$, $\beta_{i}$, and $ymax_{i}$ for every subplot $i$ and uses them as covariates in the ML models. This scenario represents an ideal situation for the researcher in which all the relevant variables are observed, but the true functional form of the heterogeneous yield response functions are not known. More realistic scenarios could be compared with this ideal scenario.

In subsequent modeling scenarios, we made the simulated experimental data more realistic by varying the degree to which the researcher understood what the needed covariates' values were those in the dataset. In the second scenario, denoted "abytt," we introduced additional variables $\theta^1$ and $\theta^2$, and the researcher's dataset contained values of $\alpha_{i}$, $\beta_{i}$, $ymax_{i}$, $\theta^1_{i}$ and $\theta^2_{i}$ for all subplots $i=1,\dots,1536$. $\theta^1$ and $\theta^2$ had no effect on yield, but were spatially correlated with $\beta_{i}$. This scenario was meant to reflect the lack of scientific consensus about which variables explain heterogeneous yield response to N, making it likely that the researcher would include irrelevant variables like $\theta^1$ and $\theta^2$ in real-world data analysis.

In the third scenario, denoted "aabbyy," the researcher's dataset did not contain values of $\alpha_{i}$, $\beta_{i}$, or $ymax_{i}$. Rather, it contained subplot-level values of permutations of $\alpha$, $\beta$, and $ymax$. Each of the three parameters was split into two parts in a spatially correlated manner, and six new covariates were created. (See Appendix C.2 for the details about this parameter-splitting process.) These were $\alpha^{1}_{i}$, $\alpha^{2}_{i}$, $\beta^{1}_{i}$, $\beta^{2}_{i}$, $ymax^{1}_{i}$, and $ymax^{2}_{i}$. The two split variables were spatially correlated to reflect that many of the soil/field characteristics used to analyze OFPE data are spatially correlated. The values of $\alpha_{i}$, $\beta_{i}$ and $ymax_{i}$ did not indicate specific soil or field characteristics, but were functions of such. In reality, scientists cannot directly observe variables that can accurately predict yield plateau level, soil N content and N uptake efficiency. Rather, they use multiple observed soil/field characteristics to explain such phenological phenomena. For example, instead of using $ymax_{i}$ directly, some topographical land features and soil properties such as elevation and soil sand content could be used as yield-limiting factors [@jiang2004effect; @kravchenko2000correlation]. In reality, scientists may include more than three variables as covariates. This scenario should reduce the accuracy of EONR modeling for all the models compared to the ideal case.

@@

The RF, BRF, and CNN methods all followed the same conceptual steps to estimate site-specific EONRs: 1) estimate the yield production functions and 2) use those estimated functions to calculate site-specific EONRs. For the given modeling scenario, RF, BRF and CNN predict yield by using all the available covariates as explanatory variables in the estimation process. The `grf` package (version 1.2.0 [@tibshirani2018package]) in R was used for RF and BRF modeling. RF and BRF from the `grf` package have several tunable hyperparameters (e.g., the minimum node size in each tree, the number of covariates used for node splitting, and parameters involving an honest tree-building process), the optimal values of which were selected by cross-validation. Individual trees in a forest were built with randomly subsampled observations, leading to a different prediction in every forest even when the same dataset was used. To reduce the variance of predictions for prediction accuracy, the number of trees (*num.trees*) was set at $2000$, a high value.
